---
title: "Online Shopping Transaction Analysis"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Incuded libraries
```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(DataExplorer)
library(arules)
library(arulesViz)
library(RColorBrewer)

```

Read file from current working directory
```{r}
online_data = read_excel("Online Retail.xlsx")

```

View data to have an idea of what it looks like
```{r}
head(online_data)
tail(online_data)
str(online_data)
summary(online_data)

```

Visualise dataset structure
```{r}
plot_str(online_data, fontSize = 40)
plot_intro(online_data)
plot_missing(online_data)

```

Change data types for the different variables and removes spaces before and after observations
```{r}

online_data$InvoiceNo = trimws(as.character(online_data$InvoiceNo))
online_data$Description = trimws(online_data$Description)


online_data = online_data%>%mutate_if(is.character, as.factor)
online_data$CustomerID = as.factor(online_data$CustomerID)
head(online_data)
str(online_data)

apply(online_data, 2, function(x) length(unique(x)))

```

Removing cancelled orders, adjusted bad debt, remove zero or negative quantities, negative unit prices and NAs
```{r}

plot(online_data$UnitPrice,ylab = "Prices", main = "Prices")
online_data[(online_data$UnitPrice <= 0),]

plot(online_data$Quantity, ylab = "Quantity", main = "Quantities Purchased")
online_data[(online_data$Quantity <= 0),]


unique(substr(online_data$InvoiceNo, 1,1))

online_data1 = filter(online_data, !(substr(online_data$InvoiceNo, 1,1) %in% c("A","C")),
                        online_data$Quantity > 0,
                        online_data$Description != "?",
                        !(online_data$Description %in% c("amazon", "AMAZON")),
                        !is.na(online_data$Description))

summary(online_data1$UnitPrice)
summary(online_data1$Quantity)

apply(online_data, 2, function(x) sum(is.na(x)))
apply(online_data1, 2, function(x) sum(is.na(x)))

dim(online_data1)
```

Aggregating Invoices and items
```{r}
online_data1_agg = split(online_data1$Description, online_data1$InvoiceNo)
head(online_data1_agg)

```

Average number of items purchased
```{r}
avg_quantity = online_data1 %>% group_by(InvoiceNo) %>% summarize(avg_basket = mean(Quantity))
head(avg_quantity)
ggplot(data = avg_quantity, aes(x = avg_quantity$avg_basket)) + geom_histogram(fill = "blue", bins = 50000) + geom_rug() + 
  coord_cartesian(xlim = c(0,80)) + labs(x = "Average Basket", y = "Frequency", title = "Average items purchased")

```

Remove duplicate items in the same invoice
```{r}
online_data1_agg1 = list()
for (i in 1:length(online_data1_agg)){
  online_data1_agg1[[i]] = unique(online_data1_agg[[i]])
}
head(online_data1_agg1)

```

Using arules to analyse transactions
```{r}
trxns = as(online_data1_agg1, "transactions")
summary(trxns)
inspect(trxns[1:10,])

```

Frequency of purchased items
```{r}
itemFrequencyPlot(trxns, support=0.05)
itemFrequencyPlot(trxns, topN=10)

```

Setting up rules
```{r}
online_basket = apriori(data = trxns, parameter = list(support = 0.005, conf = 0.75))
inspect(sort(online_basket[1:100], by = "lift"))

```

Plot Rules
```{r}
plot(online_basket, control = list(col = brewer.pal(11, "Spectral")))

```

Plot of the top 10 rules by support
```{r}
sub_rules = head(sort(online_basket, by="support"),10)
inspect(sub_rules)
plot(sub_rules, method="graph", cex = 0.5)

```

Summary
```{r}
summary(online_basket)
```

Rules based on items containing the word COFFEE
```{r}
rulesCoffee = subset(online_basket, subset= items %pin% "COFFEE")
summary(rulesCoffee)

sub_rules_coffee = head(sort(rulesCoffee, by="support"),10)
inspect(sub_rules_coffee)
plot(sub_rules_coffee, method="graph", cex = 0.5)
```

Rules based on the right hand side items having the colours Red, Black or White
```{r}
rulesColours = subset(online_basket, subset= rhs %pin% c("RED", "BLACK", "BLUE"))
summary(rulesColours)

sub_rules_colours= head(sort(rulesColours, by="support"),10)
inspect(sub_rules_colours)
plot(sub_rules_colours, method="graph", cex = 0.5)
```

